---
output: github_document
references:
- id: morris2008
  title: Time for the weather-- Translating Zulu
  author:
  - family: Morris
    given: Doug
  container-title: Weatherwise
  volume: 61
  issue: 3
  DOI: 10.3200/WEWI.61.3.32-35
  page: 32-35
  type: article-journal
  issued:
    year: 2008
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

```{r echo = FALSE}
library(countytimezones)
```

## Aims of the package

This package allows you to convert time from Coordinated Universal Time (UTC, also known as Zulu Time) to local time for US counties based on each county's FIPs code. You can use this to convert time zones in either a dataframe with all values from one county or a dataframe with values from many different time zones. 

Many observational datasets express date-times using UTC, to ensure consistency across time zones. Examples include datasets for satellite-based and / or hourly data and aviation data [@morris2008]. Other data is collected and aggregated based on local time (e.g., daily counts of health outcomes for a community). Thus, it is often helpful to be able to convert date-time values for observations from UTC to local time. This package allows you to make that conversion for any dataset where each observation can be associated with a US county Federal Information Processing Standard (FIPS) code. 

These local time conversions take into account whether or not the county was observing Daylight Savings Time at the time of the observation. This package can be used both with datasets where all observations are from the same county and for datasets where observations are associated with a number of different counties. 

## Accessing the package

This package is currently under development on GitHub. It can be installed using the `devtools` package:

```{r eval = FALSE, message = FALSE}
library(devtools)
install_github("geanders/countytimezones")
library(countytimezones)
```

## Simple example

As a simple example, if you want to convert 8:30 UTC on Jan. 1, 1999 to local time in counties throughout the US, you could run: 

```{r message = FALSE}
data(county_tzs)
example_df <- data.frame(fips = county_tzs$fips,
                 datetime = "1999-01-01 08:30")
head(example_df)

example_df <- add_local_time(df = example_df,
                             fips = example_df$fips,
                             datetime_colname = "datetime")
head(example_df)
```

You can use functions from the `choroplethr` package to map the local values:

```{r fig.width = 7, fig.height = 5, message = FALSE}
library(choroplethr)
library(ggplot2)
library(dplyr)

to_plot <- example_df %>%
  select(fips, local_time) %>%
  mutate(local_time = factor(local_time)) %>%
  dplyr::rename(region = fips, value = local_time)
a <- CountyChoropleth$new(to_plot)
a$ggplot_scale <- scale_fill_brewer(type = "qual", drop = FALSE)
a$render()
```

Here is the same map for a date-time during the summer, when many counties use Daylight Savings Time:

```{r fig.width = 7, fig.height = 5}
example_df2 <- data.frame(fips = county_tzs$fips,
                 datetime = "1999-07-01 08:30") 
example_df2 <- add_local_time(example_df2, fips = example_df2$fips,
                              datetime_colname = "datetime")

to_plot <- example_df2 %>%
  select(fips, local_time) %>%
  mutate(local_time = factor(local_time)) %>%
  dplyr::rename(region = fips, value = local_time)
a <- CountyChoropleth$new(to_plot)
a$ggplot_scale <- scale_fill_brewer(type = "qual", drop = FALSE)
a$render()
```

## More complex example

As a more complex example, the `closest_dist` data from the `hurricaneexposure` package (which can be installed from "geanders/hurricanceexposure" on GitHub) has data on the date when tropical storms were closest to US counties:

```{r fig.width = 5, fig.height = 5, warning = FALSE, message = FALSE, eval = FALSE}
# install_github("geanders/hurricaneexposure") # if you need to install the package
library(hurricaneexposure)
data(closest_dist)

floyd <- dplyr::filter(closest_dist, storm_id == "Floyd-1999")
head(floyd)

floyd <- add_local_time(floyd, fips = floyd$fips,
                        datetime_colname = "closest_date")
head(floyd)

eastern_states <- c("alabama", "arkansas", "connecticut", "delaware",
                            "district of columbia", "florida", "georgia", "illinois",
                            "indiana", "iowa", "kansas", "kentucky", "louisiana",
                            "maine", "maryland", "massachusetts", "michigan",
                            "mississippi", "missouri", "new hampshire", "new jersey",
                            "new york", "north carolina", "ohio", "oklahoma",
                            "pennsylvania", "rhode island", "south carolina",
                            "tennessee", "texas", "vermont", "virginia",
                            "west virginia", "wisconsin")

library(lubridate)
to_plot <- select(floyd, fips, closest_date) %>%
  mutate(fips = as.numeric(fips),
         closest_date = ymd_hm(closest_date)) %>%
  mutate(closest_date = format(closest_date, "%Y-%m-%d")) %>%
  dplyr::rename(region = fips, value = closest_date)
a <- CountyChoropleth$new(to_plot)
a$ggplot_scale <- scale_fill_brewer(type = "qual", drop = FALSE)
a$set_zoom(eastern_states)
a$render()

to_plot <- select(floyd, fips, local_date) %>%
  mutate(fips = as.numeric(fips))%>%
  dplyr::rename(region = fips, value = local_date)
a <- CountyChoropleth$new(to_plot)
a$ggplot_scale <- scale_fill_brewer(type = "qual", drop = FALSE)
a$set_zoom(eastern_states)
a$render()
```

```{r warning = FALSE, message = FALSE, fig.width = 6, fig.height = 4, eval = FALSE}
to_plot <- select(floyd, fips, closest_date) %>%
  filter(substring(fips, 1, 2) %in% c("47")) %>%
  mutate(fips = as.numeric(fips), 
         closest_date = ymd_hm(closest_date)) %>%
  mutate(closest_date = format(closest_date, "%Y-%m-%d %H")) %>%
  mutate(closest_date = factor(closest_date)) %>%
  dplyr::rename(region = fips, value = closest_date) %>%
  filter(value %in% levels(value)[2:9]) %>%
  mutate(value = factor(value))

a <- CountyChoropleth$new(to_plot)
a$ggplot_scale <- scale_fill_brewer(type = "qual", drop = FALSE)
a$set_zoom(c("tennessee"))
a$render()

to_plot <- select(floyd, fips, local_time) %>%
  filter(substring(fips, 1, 2) %in% c("47")) %>%
  mutate(fips = as.numeric(fips), 
         local_time = ymd_hm(local_time)) %>%
  mutate(local_time = format(local_time, "%Y-%m-%d %H")) %>%
  mutate(local_time = factor(local_time)) %>%
  dplyr::rename(region = fips, value = local_time) %>%
  filter(value %in% levels(value)[2:9]) %>%
  mutate(value = factor(value))

a <- CountyChoropleth$new(to_plot)
a$ggplot_scale <- scale_fill_brewer(type = "qual", drop = FALSE)
a$set_zoom(c("tennessee"))
a$render()
```

## References
